Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-19 document/ms.tex
expanding on the extrapolation section

\item the observational uncertainties on the physical parameters are
non-negligible but the amplitudes are known (the catalog provides
probabilistic constraints on the parameters),

@@

\section{Model generalities}

@@

As well as inferring the ocurrence distribution of exoplanets, this dataset
can also be used to constrain the rate of Earth analogs.
Even though no true analog has been observed, we can make some assumptions
about the distribution (as described below) and extrapolate to the location of
the Earth.
Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
location of Earth}
\begin{eqnarray}
\gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
&=&
\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
\end{eqnarray}
This is also the quantity that \citet{petigura} estimated but in different
units.
That is, \gammaearth\ is the average number of exoplanets around a
``Sun-like'' star---per natural logarithmic bins in period and
radius---evaluated at the period and radius of Earth.

>>>>

\item the observational uncertainties on the physical parameters are
non-negligible but known (the catalog provides probabilistic constraints on
the parameters),

@@

\section{Model generalities}
\sectlabel{model}

@@

As well as inferring the ocurrence distribution of exoplanets, this dataset
can also be used to constrain the rate of Earth analogs.
Explicitly, we constrain the rate of exoplanets \emph{evaluated at the
location of Earth}
\begin{eqnarray}
\gammaearth &=& \rate (\ln\period_\oplus,\,\ln\radius_\oplus) \\
&=&
\left.\frac{\dd N}{\dd\ln\period\,\dd\ln\radius}\right|
_{\radius=\radius_\oplus,\,\period=\period_\oplus}\quad.
\end{eqnarray}
This is also the quantity that \citet{petigura} estimated but in different
units.
That is, \gammaearth\ is the average number of exoplanets around a
``Sun-like'' star---per natural logarithmic bins in period and
radius---evaluated at the period and radius of Earth.

Since no Earth analogs have been found any constraints on this rate must be
extrapolated from the existing observations.
This extrapolation is generally done by assuming a functional form for the
occurence rate function, constraining it using the observed candidates and
extrapolating.
All published extrapolations are based on rigid models like a power law
(\citealt{catanzarite, traub}) or linear in the cumulative period distribution
(\citealt{petigura}).
All of these results are sensitive to this choice of function and to the other
implicit decisions (for example the size of radius bin that is considered
Earth-like by \citealt{petigura}).
We weaken the assumptions necessary for extrapolation by only assuming that
the distribution is smooth using the Gaussian process prior described in
\sect{model}.
This model will, of course, break down if there is a sharp feature in the
occurence rate function right at the location of Earth but otherwise this is
the most conservative extrapolation method published up to this point.

The results of our extrapolation method, when applied to the synthetic
catalogs described in the previous section, are shown in \fig{smooth-rate} and
\fig{simulation-rate}.
We implemented the extrapolation method from \citet{petigura} that assumes a
linear relation in the cumulative period distribution for small planets and
compared these results---converting to the correct units---to our method and
the truth.
In both cases, our method returns a less precise but more accurate result for
the occurence rate and the error bars given by the functional extrapolation
are overly optimistic.
